---
title: "build cox model for cicu and msicu data"
author: "JunmeiW"
date: "2019年5月1日"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "F:/.../2019-2plus/R_code2019/data/")
```

##1. set working directory
```{r}
library(readr)
setwd("F:/.../2019-2plus/R_code2019/data/")
```


## 1.get all data:time-series

```{r}
library(readr)
AA_com_info <- read_csv("data/aa_com_info2.csv")
```

## 1.1 get CICU DATA AND MSICU DATA seperately
```{r}
library(sqldf)
# unique(AA_com_info$unittype)
cicu_7day = sqldf("select * from AA_com_info where lower(unittype) like 'c%'",row.names = TRUE)
sicu_7day = sqldf("select * from AA_com_info where lower(unittype) like 's%' or 
                  lower(unittype) like 'm%'",row.names = TRUE)
```


```{r}
library(survival)
library(KMsurv)
library(RColorBrewer)
library(survminer)
```


## 2. four bin KM Curve- cicu_7day

```{r fig.height=4, fig.width=5}
kmc.7day.1 <- survfit(Surv(tstart, tstop, status == 1) ~ 1, data = cicu_7day, 
                     id= cicu_7day$icustay_id, conf.type = "log-log" )
plot(kmc.7day.1, fun='event') 
kmc.7day.4 <- survfit(Surv(tstart, tstop, status == 1) ~ factor(four_bin), data = cicu_7day, 
                      id= cicu_7day$icustay_id, conf.type = "log-log" )

pdf("data/update0405/cicu_KM.pdf")
plot(kmc.7day.4, fun='event', xlab = "Time(hours)", ylab="Probabilities" , 
     ylim = c(0.0,0.165), 
     col = brewer.pal(4,"Dark2"))
legend("topleft",inset = .0, title = "Potassium value",col = brewer.pal(4,"Dark2"),  
       #brewer.pal(4,"Set3") Dark2
        c("≥4.0 and ≤5.0 mEq/L","≥3.5 and <4.0 mEq/L","≥3.0 and <3.5 mEq/L","<3.0 mEq/L"),lty = 1)
dev.off()
```

### 2.1 
```{r}
library(ggplot2)
cph  <- coxph(Surv(tstart, tstop, status == 1) ~ strata(four_bin), data = cicu_7day
                     )
base <- basehaz(cph)
baseline_gg <- ggplot(base, aes(x=time)) +
  geom_step(aes(y=hazard, group=strata)) +
    ylab(expression(hat(Lambda)(t))) + xlab("t")
baseline_gg + aes(col=strata)
```


###  Using appropriate  cols
```{r}
library(dplyr)
cicu_7day = subset(cicu_7day, select = c( icustay_id, tstart,tstop,status,
                                         age,gender, apache_iv, charlson_score,
                                         four_bin,vaso_inotro,furosemide))

sicu_7day = subset(sicu_7day, select = c(icustay_id, tstart,tstop,status,
                                         age,gender, apache_iv, charlson_score,
                                         four_bin,vaso_inotro,furosemide))
```


--------------------- 2. have no covariates(unadjusted model)------------------

## 2.1 Unadjusted cox model for cicu data
```{r}
unadj_c1 <- coxph(Surv(tstart,tstop, status) ~ (four_bin),
                 data = cicu_7day ) 
unadj_cz<- cox.zph(unadj_c1, transform = "identity") 
print(unadj_cz)
# summary(unadj_c1)
```

```{r}
plot(unadj_cz[1], df=2, xlim=c(0,168),ylim=c(-1,0.2))
abline(h = coef(unadj_c)[1], lty = 2, col = "red")
```


## dour bin(==2) violates the PH, including time interaction
```{r}
time_unadjc <- update(unadj_c1,.~.+model.matrix(~four_bin)[,c(2)]:(tstart)
                      )    
summary(time_unadjc)
```



## 2.2 Unadjusted cox model for cicu data
```{r}
unadj_s <- coxph(Surv(tstart,tstop, status ) ~ (four_bin), data = sicu_7day ) 

unadj_sz<- cox.zph(unadj_s, transform = "identity") 
print(unadj_sz)
```


```{r}
time_unadjs <- update(unadj_s,.~.+model.matrix(~four_bin)[,c(3)]:(tstart)
                      +model.matrix(~four_bin)[,c(4)]:(tstart))
                       
summary(time_unadjs)
```

----------------------------3. Building Multiple variates COX model -------------------------------------------------------

- **convert age into 4 categories**
```{r}
# summary(unique(cicu_7day$age))
ctr_apa = unique(cicu_7day[,c(1,5)])
summary(ctr_apa)
cicu_7day$age_cat <- 0
cicu_7day$age_cat[cicu_7day$age <  54.00] <- 1
cicu_7day$age_cat[(cicu_7day$age >= 54) & (cicu_7day$age <  65.00 )] <- 2
cicu_7day$age_cat[(cicu_7day$age >= 65.00) & (cicu_7day$age < 76.00)] <- 3
cicu_7day$age_cat[(cicu_7day$age >= 76.00)] <- 4
cicu_7day$age_cat <- as.factor(cicu_7day$age_cat)
```


- **convert Apache IV score into 4 categories**
```{r}
ctr_apa = unique(cicu_7day[,c(1,7)])
summary(ctr_apa)
cicu_7day$apache_cat <- 0
cicu_7day$apache_cat[cicu_7day$apache_iv <  35.00] <- 1
cicu_7day$apache_cat[(cicu_7day$apache_iv >= 35.00) & (cicu_7day$apache_iv <  47.00 )] <- 2
cicu_7day$apache_cat[(cicu_7day$apache_iv >= 47.00) & (cicu_7day$apache_iv < 61)] <- 3
cicu_7day$apache_cat[(cicu_7day$apache_iv >= 61.00)] <- 4
cicu_7day$apache_cat <- as.factor(cicu_7day$apache_cat)
```

```{r}
cicu_cm.cat <- coxph(Surv(tstart,tstop, status ) ~ (four_bin) + age_cat + gender + apache_cat + 
                    charlson_score +furosemide + vaso_inotro, data = cicu_7day ) 
cicu_cmz.cat<- cox.zph(cicu_cm.cat, transform = "identity") 
print(cicu_cmz.cat)
```
- ** plot beta(t) for variable**
```{r}
plot(cicu_cmz.cat[1], df=2, xlim=c(0,168))
abline(h = coef(cicu_cm.cat)[1], lty = 2, col = "red")
plot(cicu_cmz.cat[2], df=2, xlim=c(0,168))
abline(h = coef(cicu_cm.cat)[2], lty = 2, col = "red")
```


```{r}
plot(cicu_cmz.cat[10], df=2, xlim=c(0,168))
abline(h = coef(cicu_cm.cat)[10], lty = 2, col = "red")

plot(cicu_cmz.cat[11], df=2, xlim=c(0,168))
abline(h = coef(cicu_cm.cat)[11], lty = 2, col = "red")

plot(cicu_cmz.cat[13], df=2, xlim=c(0,168), ylim=c(1,2))
abline(h = coef(cicu_cm.cat)[13], lty = 2, col = "red")
```


### including time-variable interaction term into the model 
```{r}
time_cicu <- update(cicu_cm.cat,.~.+model.matrix(~four_bin)[,c(2)]:(tstart)+ # log(t+1)
                      model.matrix(~four_bin)[,c(3)]:(tstart)+
                          model.matrix(~age_cat)[,c(3)]:(tstart)+
                       model.matrix(~age_cat)[,c(4)]:(tstart)+
                           model.matrix(~vaso_inotro)[,c(-1)]:log(tstart+1)+
                      charlson_score:tstart
                             )

summary(time_cicu)
```


```{r}
# summary(tcicu_7day.0)
library(tab)

tc = summary(time_cicu)

# Prepare the columns
HR <- round(exp(coef(time_cicu)), 3)
CI <- round(exp(confint(time_cicu)), 3)
P_value <- tc$coefficients[, c(5)]

P_value <- round(P_value, 3)

# Names the columns of CI
colnames(CI) <- c("Lower Bound", "Higher Bound")

# Bind columns together as dataset
tablec <- as.data.frame(cbind(HR, CI,P_value))
tablec
```

### Write the table results into documment
```{r}
library(ReporteRs)
library(magrittr)

# The script
docx( ) %>% 
     addFlexTable(tablec %>%
               FlexTable(header.cell.props = cellProperties( background.color = "#003366"),
                    header.text.props = textBold(color = "white"),
                    add.rownames = TRUE ) %>%
               setZebraStyle(odd = "#DDDDDD", even = "#FFFFFF")) %>%
     writeDoc(file = "data/update0405/table_cicu3.docx")

# write.csv(c3,"data/finalcox/cicu_coxfinal.csv")
```

------------------------------------ **Plot forest plot(Hazard ratio)** --------------------------------

### Prepare data
- "≥4.0 and ≤5.0 mEq/L","≥3.5 and <4.0 mEq/L","≥3.0 and <3.5 mEq/L","<3.0 mEq/L")
```{r}
forst_cicu <-  structure(list(
    HR = c(NA,NA, tablec$HR[1:3],NA,tablec$HR[4:6],tablec$HR[7],NA,tablec$HR[8:10],tablec$HR[11:13]),
    lower  = c(NA,NA, tablec$`Lower Bound`[1:3],NA,tablec$`Lower Bound`[4:6],tablec$`Lower Bound`[7],NA,
    tablec$`Lower Bound`[8:10],tablec$`Lower Bound`[11:13]),
    upper = c(NA,NA, tablec$`Higher Bound`[1:3],NA,tablec$`Higher Bound`[4:6],tablec$`Higher Bound`[7],NA,
    tablec$`Higher Bound`[8:10],tablec$`Higher Bound`[11:13])
    ),
    .Names = c("HR","Lower_CI", "Upper_CI"), 
    row.names = c(NA,"Potassium Category (≥4 and ≤5 mEq/L as reference)","≥3.5 and <4.0 mEq/L",
"3.0 and <3.5 mEq/L",
"3.0 mEq/L","Age(17~54 as reference)","54~65","65~76","76~91.4","Gender, male",
"APACHE IV(< 35 as reference)",
"35~47",
"47~61",
"61~211",
"Charlson score",
"furosemide usage (timeseries)",
"vasopressor or inotropic medication usage(timeseries)"),
    class = "data.frame")
    
    
tabletext <- cbind(
  c(NA,
"Potassium Category (≥4 and ≤5  mEq/L as reference)",
"≥3.5 and <4.0 mEq/L",
"3.0 and <3.5 mEq/L",
"3.0 mEq/L","Age(17~54 as reference)","54~65","65~76","76~91.4","Gender, male",
"APACHE IV (< 35 as reference)",
"35~47",
"47~61",
"61~211",
"Charlson score",
"furosemide usage (timeseries)",
"vasopressor or inotropic medication usage(timeseries)"
),
  c("Lower_CI",NA, tablec$`Lower Bound`[1:3],NA,tablec$`Lower Bound`[4:6],tablec$`Lower Bound`[7],NA,
    tablec$`Lower Bound`[8:10],tablec$`Lower Bound`[11:13]),
  c("Upper_CI",NA,tablec$`Higher Bound`[1:3],NA,tablec$`Higher Bound`[4:6],tablec$`Higher Bound`[7],NA,
    tablec$`Higher Bound`[8:10],tablec$`Higher Bound`[11:13]),
  c("Hazard Ratio", NA,tablec$HR[1:3],NA,tablec$HR[4:6],tablec$HR[7],NA,
    tablec$HR[8:10],tablec$HR[11:13])
)
```


### plot forestplot for CICU model(Just potassium bin)
```{r}
library(haven)
library(forestplot)
forst_cicu <- forst_cicu[0:5,]
tabletext <- tabletext[0:5,]
# tablec$P_value[tablec$P_value == '0.000'] <- '<0.001'

pdf("data/update0405/results0409/cicu7day.pdf")
forestplot(tabletext[,c(1,4)],
                       mean = forst_cicu$HR, 
                       lower = forst_cicu$Lower_CI,
                       upper = forst_cicu$Upper_CI,
                       # new_page = TRUE,
           zero = 1,
                       boxsize= 0.1,
                       lineheight = "auto",linemargin = unit(5,"mm"),
                       xticks = seq(0,2.5,0.5),
                       # xlab = "Hazard Ratio with 95%CI",
                       ci.vertices =T,
                       txt_gp = fpTxtGp(ticks = gpar(cex = 0.9),
                                        label = gpar(fontfamily = "Times"), cex = 0.8),
                       # is.summary=c(TRUE,TRUE,rep(TRUE,8),TRUE),
                       col = fpColors(box=c("royalblue"),
                          line=c("darkblue"))
                       
                       )
dev.off()
```

## SICU data
```{r}
# summary(unique(cicu_7day$age))
ctr_apa = unique(sicu_7day[,c(1,5)])
summary(ctr_apa)
sicu_7day$age_cat <- 0
sicu_7day$age_cat[sicu_7day$age <  50.00] <- 1
sicu_7day$age_cat[(sicu_7day$age >= 50) & (sicu_7day$age <  63.00 )] <- 2
sicu_7day$age_cat[(sicu_7day$age >= 63.00) & (sicu_7day$age < 75.00)] <- 3
sicu_7day$age_cat[(sicu_7day$age >= 75.00)] <- 4
sicu_7day$age_cat <- as.factor(sicu_7day$age_cat)
```





```{r}
ctr_apa = unique(sicu_7day[,c(1,7)])
summary(ctr_apa)
sicu_7day$apache_cat <- 0
sicu_7day$apache_cat[sicu_7day$apache_iv <  35.00] <- 1
sicu_7day$apache_cat[(sicu_7day$apache_iv >= 35.00) & (sicu_7day$apache_iv <  48.00 )] <- 2
sicu_7day$apache_cat[(sicu_7day$apache_iv >= 48.00) & (sicu_7day$apache_iv < 64)] <- 3
sicu_7day$apache_cat[(sicu_7day$apache_iv >= 64.00)] <- 4
sicu_7day$apache_cat <- as.factor(sicu_7day$apache_cat)
```


update model 
```{r}
sicu_cm.cat <- coxph(Surv(tstart,tstop, status ) ~ (four_bin) + age_cat + gender + apache_cat + 
                    charlson_score +furosemide + vaso_inotro, data = sicu_7day ) 
sicu_cmz.cat<- cox.zph(sicu_cm.cat, transform = "identity") 
print(sicu_cmz.cat)

# summary(sicu_cm.cat)
```

```{r}
plot(sicu_cmz.cat[4], df=2, xlim=c(0,168))
abline(h = coef(sicu_cm.cat)[4], lty = 2, col = "red")

plot(sicu_cmz.cat[5], df=2, xlim=c(0,168))
abline(h = coef(sicu_cm.cat)[5], lty = 2, col = "red")

plot(sicu_cmz.cat[6], df=2, xlim=c(0,168))
abline(h = coef(sicu_cm.cat)[6], lty = 2, col = "red")
```

```{r}
time_sicu <- update(sicu_cm.cat,.~.+model.matrix(~four_bin)[,c(3)]:(tstart)+
                      model.matrix(~four_bin)[,c(4)]:(tstart)+
                      model.matrix(~age_cat)[,c(4)]:(tstart)+
                       model.matrix(~apache_cat)[,c(4)]:(tstart)+
                           model.matrix(~vaso_inotro)[,c(-1)]:(tstart)+
                      charlson_score:tstart
                             )
summary(time_sicu)
```


```{r}
library(ReporteRs)
library(magrittr)
sicu_m = summary(time_sicu)
HR <- round(exp(coef(time_sicu)), 3)
CI <- round(exp(confint(time_sicu)), 3)
P_value <- round(sicu_m$coefficients[, c(5)],3)
# Names the columns of CI
colnames(CI) <- c("Lower Bound", "Higher Bound")

# Bind columns together as dataset
table_sicu <- as.data.frame(cbind(HR,CI,P_value))
table_sicu 

# The script
docx( ) %>% 
     addFlexTable(table_sicu %>%
               FlexTable(header.cell.props = cellProperties( background.color = "#003366"),
                    header.text.props = textBold(color = "white"),
                    add.rownames = TRUE ) %>%
               setZebraStyle(odd = "#DDDDDD", even = "#FFFFFF")) %>%
     writeDoc(file = "data/update0405/table_sicu.docx")


```


```{r}
forst_sicu <-  structure(list(
    HR = c(NA,NA, table_sicu$HR[1:3],NA,table_sicu$HR[4:6],table_sicu$HR[7],NA,
           table_sicu$HR[8:10],table_sicu$HR[11:13]),
    lower  = c(NA,NA, table_sicu$`Lower Bound`[1:3],NA,table_sicu$`Lower Bound`[4:6],
               table_sicu$`Lower Bound`[7],NA,table_sicu$`Lower Bound`[8:10],
               table_sicu$`Lower Bound`[11:13]),
    upper = c(NA,NA, table_sicu$`Higher Bound`[1:3],NA,table_sicu$`Higher Bound`[4:6],
              table_sicu$`Higher Bound`[7],NA,
    table_sicu$`Higher Bound`[8:10],table_sicu$`Higher Bound`[11:13])),
    .Names = c("HR","Lower_CI", "Upper_CI"), 
    row.names = c(NA,"Potassium Category (≥4 and ≤5 mEq/L as reference)","3.5~4.0",
"3.0~3.5",
"<3","Age(17~50 as reference)","50~63","63~75","75~91.4","Gender, male",
"APACHE IV(< 35 as reference)",
"35~48",
"48~64",
"64~199",
"Charlson score",
"furosemide usage (timeseries)",
"vasopressor or inotropic medication usage(timeseries)"),
    class = "data.frame")
    
    
tabletext <- cbind(
  c(NA,
"Potassium Category (≥4 and ≤5 mEq/L as reference)",
"≥3.5 and <4.0 mEq/L",
"≥3.0 and <3.5 mEq/L",
"<3.0 mEq/L","Age(17~50 as reference)","50~63","63~75","75~91.4","Gender, male",
"APACHE IV (< 35 as reference)",
"35~48",
"48~64",
"64~199",
"Charlson score",
"furosemide usage (timeseries)",
"vasopressor or inotropic medication usage(timeseries)"
),
  c("Lower_CI",NA, table_sicu$`Lower Bound`[1:3],NA,
    table_sicu$`Lower Bound`[4:6],table_sicu$`Lower Bound`[7],NA,
    table_sicu$`Lower Bound`[8:10],table_sicu$`Lower Bound`[11:13]),
  c("Upper_CI",NA,table_sicu$`Higher Bound`[1:3],NA,table_sicu$`Higher Bound`[4:6],
    table_sicu$`Higher Bound`[7],NA,
    table_sicu$`Higher Bound`[8:10],table_sicu$`Higher Bound`[11:13]),
  c("Hazard Ratio", NA,table_sicu$HR[1:3],NA,table_sicu$HR[4:6],table_sicu$HR[7],NA,
    table_sicu$HR[8:10],table_sicu$HR[11:13])
)
```

plot forestplot for SICU hazard ratio
```{r}
library(haven)
library(forestplot)

forst_sicu <- forst_sicu[0:5,]
tabletext <- tabletext[0:5,]
pdf("data/update0405/results0409/sicu7day.pdf")
forestplot(tabletext[,c(1,4)],
                       mean = forst_sicu$HR, 
                       lower = forst_sicu$Lower_CI,
                       upper = forst_sicu$Upper_CI,
                       new_page = FALSE,
                       boxsize= 0.1,
                       zero = 1,
                       lineheight = "auto",linemargin = unit(5,"mm"),
                       xticks = seq(0,2.01,0.5),
                       # xlab = "Hazard Ratio with 95%CI",
                       ci.vertices =T,
                       txt_gp = fpTxtGp(ticks = gpar(cex = 0.9),
                                        label = gpar(fontfamily = "Times"), cex = 0.8),
                       # is.summary=c(TRUE,TRUE,rep(TRUE,8),TRUE),
                       col = fpColors(box=c("royalblue"),
                          line=c("darkblue"))
                       
                       )
dev.off()
```
